---
title: "lmForc"
author:
  - name: "Nelson Rayl"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lmForc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(lmForc)
```

## Sections
#### [Forecast Class]
#### [is_forc]
#### [oos_realized_forc]
#### [oos_lag_forc]
#### [oos_vintage_forc]
#### [conditional_forc]
#### [mse_weighted_forc]

## lmForc Motivation

Linear forecasting models have two main features: simplicity and 
interpretability. These features allow the performance of linear forecasting 
models to be tested in multiple ways, both in-sample and out-out-sample. The
*lmForc* package brings these performance tests to the R language with a
philosophy that compliments the features of linear models: One, the syntax for
creating linear models is simple and the syntax for testing them should be 
equally simple. Two, performance tests should be realistic and when possible
replicate what it would have been like to forecast in real time.

## Forecast Class

At the heart of the *lmForc* package is the `Forecast` class. Base R does not
provide a good format for working with forecasts, so *lmForc* addresses this by
introducing a new class for storing forecasts that is simple and rigorous. The
`Forecast` class is paramount to the *lmForc* philosophy of simple syntax and
realistic tests. `Forecast` is an S4 class that contains equal length vectors 
with the following data: 

* `origin` \ \ \ The time when the forecast was made.
* `future` \ \ \ The time that is being forecasted.
* `forecast` The forecast itself.
* `realized` \ If available, the realized value at the time being forecasted.

The `Forecast` class also includes an additional length-one slot `h_ahead` for
representing how many periods ahead are being forecasted. This slot is optional,
but becomes useful for documentation and performing out-of-sample forecast 
tests.

We demonstrate the `Forecast` class by constructing a simple `Forecast` object. 
This forecast contains four observations at the quarterly frequency.

```{r}
my_forecast <- Forecast(
   origin   = as.Date(c("2010-03-31", "2010-06-30", "2010-09-30", "2010-12-31")),
   future   = as.Date(c("2011-03-31", "2011-06-30", "2011-09-30", "2011-12-31")),
   forecast = c(4.21, 4.27, 5.32, 5.11),
   realized = c(4.40, 4.45, 4.87, 4.77),
   h_ahead  = 4L
)
```

Note that we have chosen Date objects at the quarterly frequency for our
`origin` and `future` slots. This forecast is for four quarters ahead, so we
fill in the `h_ahead` slot with the integer four. 

The `origin`, `future`, and `h_ahead` slots can be filled with any values. This
is where the general nature of the `Forecast` class shines. In the `origin` and
`future` slots we could use dates at a daily or yearly frequency, the `POSIXct`
class to include minutes and seconds, or integers to represent discrete periods.
We can also store different types of forecasts. In the example above, we have
forecasts made at different `origin` times for a constant four quarter ahead 
forecast horizon. We could also store a forecast made at a single `origin` time
for a horizon of `future` times by setting all of the `origin` values to one
time and using `future` to represent the horizon of times that we are
forecasting. In this case the `h_ahead` slot becomes irrelevant and it is left
as `NULL`. The flexibility of these slots allows us to represent any type of
numeric forecast. 

The `forecast` and `realized` slots take numeric vectors. In the `forecast` slot
we see the forecast that was made at each `origin` time and in the `realized`
slot we see the true value that was realized at each `future` time. The realized 
values may not exist yet, so this slot may be partially populated or not populated
at all. If the `realized` slot is set to `NULL` then it will be
populated with a vector of `NA` values. 

The `Forecast` class strikes a balance between simplicity and rigor. It is 
simple enough to store any numeric forecast, but it is rigorous enough to create
a useful data structure. For example, we can quickly calculate the MSE or RMSE
of the `Forecast` object we created above with only one argument.

```{r}
mse(my_forecast)
rmse(my_forecast)
```

Because the `forecast` and `realized` slots must be numeric vectors, and all 
slots must be of the same length, we do not have to worry about input validation 
or coercing multiple vectors to the correct format when calculating MSE or RMSE.

MSE is calculated as:

$$
\hspace{8.45cm} n = \text{forecast vector length}\\[.20cm]
MSE = \frac{1}{n} \sum_{i=1}^{n}{(Y_i - \hat{Y_i})^2} 
\qquad \text{where} \quad Y_i = \text{realized values} \\
\hspace{7.3cm} \hat{Y_i} = \text{forecast values}
$$

RMSE is calculated as:
$$ RMSE = \sqrt{MSE}  $$

Note that these equations require two inputs, but because both inputs are 
already stored in the `Forecast` object we only need to pass one argument to the
`mse()` and `rmse()` functions. Calculating forecast accuracy is a simple use 
case of the `Forecast` class. More complex use cases exist in the *lmForc*
package, where many of the functions require inputs to be of the `Forecast`
class.  When weighting multiple forecasts or testing a linear model that is
conditional on another forecast, the consistent structure of the class results 
in simple functions that execute correctly, no matter the type of forecast 
passed to the function. Furthermore, all *lmForc* functions return objects of
the `Forecast` class which creates synergy between functions. One
can take two linear models, test their performance out-of-sample with the
`oos_realized_forc()` function which returns `Forecast` objects, and then pass 
these two objects to the `mse_weighted_forc()` function to find the weighted 
out-of-sample performance of both models.

One fear may be that the novel `Forecast` class will not play well with 
functions and packages that already exist in the R language. The *lmForc*
package provides methods for accessing all of the vectors stored in a `Forecast`
object as well as the `collect()` function which returns one or multiple 
`Forecast` objects as a data frame. With these methods, one can easily pass the
data in a `Forecast` object to other functions.

```{r}
collect(my_forecast)

origin(my_forecast)

future(my_forecast)

forc(my_forecast)

realized(my_forecast)
```

## Example Dataset

Examples throughout the rest of the vignette will use a stylized dataset with a
`date` column of ten quarterly dates, a dependent variable `y`, and two
independent variables `x1` and `x2`. Equations are also written in terms of the 
variables `y`, `x1`, and `x2`.

```{r}
date <- as.Date(c("2010-03-31", "2010-06-30", "2010-09-30", "2010-12-31",
                  "2011-03-31", "2011-06-30", "2011-09-30", "2011-12-31", 
                  "2012-03-31", "2012-06-30"))

y  <- c(1.09, 1.71, 1.09, 2.46, 1.78, 1.35, 2.89, 2.11, 2.97, 0.99)

x1 <- c(4.22, 3.86, 4.27, 5.60, 5.11, 4.31, 4.92, 5.80, 6.30, 4.17)

x2  <- c(10.03, 10.49, 10.85, 10.47, 9.09, 10.91, 8.68, 9.91, 7.87, 6.63)

data <- data.frame(date, y, x1, x2)

head(data)
```

## is_forc

The `is_forc()` function produces an in-sample forecast based on a linear model.
The function takes a linear model call and an optional vector of time data
associated with the linear model as inputs. The linear model is estimated once over the
entire sample period and the coefficients are multiplied by the realized values 
in each period of the sample. This is functionally identical to the `predict()` 
function in the *stats* package, but it returns a `Forecast` object instead of a
numeric vector.

For all observations *i* in the sample, coefficients are estimated as:

$$ 
Y_i = \beta_0 + \beta_1 X1_i + \beta_2 X2_i + \epsilon_i 
\qquad \text{for all } i
$$
And forecasts are estimated as:

$$ 
forecast_i = \beta_0 + \beta_1 X1_i + \beta_2 X2_i \qquad
$$

```{r}
is_forc(
  lm_call  = lm(y ~ x1 + x2, data),
  time_vec = data$date
)
```

Note that because we are creating an in-sample forecast, `h_ahead` is set to 0 
and the `origin` time equals the `future` time. This test evaluates how well a
linear forecast model fits the historical data.

## oos_realized_forc

The `oos_realized_forc()` function produces an *h* period ahead out-of-sample
forecast that is conditioned on realized values. The function takes a linear 
model call, an integer number of periods ahead to forecast, a period to end the
initial coefficient estimation and begin forecasting, and an optional vector of
time data associated with the linear model. The linear model is originally
estimated with data up to `estimation_end`. Coefficients are multiplied by
realized values of the covariates `h_ahead` periods ahead to create an `h_ahead`
period ahead forecast. This process is iteratively repeated for each period
after `estimation_end` with coefficients updating in each period as more
information would have become available to the forecaster.

In the sample *i*, for each period *p* greater than or equal to
`estimation_end`, coefficients are updated as:

$$
Y_i = \beta_0 + \beta_1 X1_i + \beta_2 X2_i + \epsilon_i 
\qquad \text{for all} \quad i \leq \text{p}
$$
And `h_ahead` *h* forecasts are estimated as:

$$
forecast_{p+h} = \beta_0 + \beta_1 X1_{p+h} + \beta_2 X2_{p+h} \qquad
$$

```{r}
oos_realized_forc(
  lm_call = lm(y ~ x1 + x2, data),
  h_ahead = 2L,
  estimation_end = as.Date("2011-03-31"),
  time_vec = data$date
)
```

Note that the `oos_realized_forc` function returns an out-of-sample forecast 
that conditions on realized values that **would not** have been available to the 
forecaster at the forecast origin. This test evaluates the performance of a 
linear forecast model had it been conditioned on perfect information.

## oos_lag_forc

The `oos_lag_forc()` function produces an *h* period ahead out-of-sample 
forecast conditioned on present period values. The function takes a linear model
call, an integer number of periods ahead to forecast, a period to end the
initial coefficient estimation and begin forecasting, and an optional vector of 
time data associated with the linear model. The linear model data is lagged by
`h_ahead` periods and the linear model is re-estimated with data up to
`estimation_end` to create a lagged linear model. Coefficients are multiplied by 
present period realized values of the covariates to create a forecast for
`h_ahead` periods ahead. This process is iteratively repeated for each period
after `estimation_end` with coefficients updating in each period.

In the sample *i*, for each period *p* greater than or equal to
`estimation_end`, coefficients are updated as:

$$
Y_i = \beta_0 + \beta_1 X1_{i-h} + \beta_2 X2_{i-h} + \epsilon_i 
\qquad \text{for all} \quad i \leq \text{p}
$$
Where *h* is `h_ahead`. `h_ahead` forecasts are estimated as:

$$
forecast_{p+h} = \beta_0 + \beta_1 X1_{p} + \beta_2 X2_{p} \qquad
$$

```{r}
oos_lag_forc(
  lm_call = lm(y ~ x1 + x2, data),
  h_ahead = 2L,
  estimation_end = as.Date("2011-03-31"),
  time_vec = data$date
)
```

This test evaluates the performance of a lagged linear model had it been
conditioned on present values that **would** have been available to the
forecaster at the forecast origin. This is in contrast to conditioning on 
realized values or a forecast of the covariates.

## oos_vintage_forc

The `oos_vintage_forc()` function produces an out-of-sample forecast conditioned
on *h* period ahead forecasts of the linear model covariates. The function takes
a linear model call, a vector of time data associated with the linear model, and
a vintage forecast for each covariate in the linear model. For each period in 
the vintage forecasts, coefficients are updated based on information that would 
have been available to the forecaster at the forecast origin. Coefficients are
then multiplied by vintage forecast values.

In the sample *i*, for each period *p* in the vintage forecasts *VF1* and *VF2*,
coefficients are updated as:

$$
Y_i = \beta_0 + \beta_1 X1_i + \beta_2 X2_i + \epsilon_i 
\qquad \text{for all} \quad i \leq \text{p}
$$

And `h_ahead` *h* forecasts are estimated as:

$$
forecast_{p+h} = \beta_0 + \beta_1 VF1_p + \beta_2 VF2_p \qquad
$$

We introduce stylized vintage forecasts of X1 and X2 to demonstrate the 
`oos_vintage_forc()` function. Using four quarter ahead forecasts of the 
covariates X1 and X2, we create an out-of-sample forecast based on the
coefficients and covariate forecasts that the forecaster would have used in 
each period.

```{r}
x1_forecast_vintage <- Forecast(
   origin   = as.Date(c("2010-09-30", "2010-12-31", "2011-03-31", "2011-06-30")),
   future   = as.Date(c("2011-09-30", "2011-12-31", "2012-03-31", "2012-06-30")),
   forecast = c(6.30, 4.17, 5.30, 4.84),
   realized = c(4.92, 5.80, 6.30, 4.17),
   h_ahead  = 4L
)

x2_forecast_vintage <- Forecast(
   origin   = as.Date(c("2010-09-30", "2010-12-31", "2011-03-31", "2011-06-30")),
   future   = as.Date(c("2011-09-30", "2011-12-31", "2012-03-31", "2012-06-30")),
   forecast = c(7.32, 6.88, 6.82, 6.95),
   realized = c(8.68, 9.91, 7.87, 6.63),
   h_ahead  = 4L
)

oos_vintage_forc(
  lm_call = lm(y ~ x1 + x2, data),
  time_vec = data$date,
  x1_forecast_vintage, x2_forecast_vintage
)
```

This test replicates the forecasts that a linear model conditional on forecasts 
of covariates would have produced in real time. Here we see the strength of the
`Forecast` class. Because the vintage forecasts of X1 and X2 share the same data
structure, we can calculate a forecast that is conditional on these objects 
without fearing inconsistency across inputs.

## conditional_forc

The `conditional_forc()` function produces a forecast conditioned on forecasts
of the linear model covariates. The function takes a linear model call, a vector
of time data associated with the linear model, and a forecast for each covariate
in the linear model. The linear model is estimated once over the entire sample
period and the coefficients are multiplied by the forecasts of each covariate. 

For all observations *i* in the sample, coefficients are estimated as:

$$
Y_i = \beta_0 + \beta_1 X1_i + \beta_2 X2_i + \epsilon_i 
\qquad \text{for all } i
$$

And for all periods *p* in the covariate forecasts *F1* and *F2*, forecasts are estimated as:

$$ 
forecast_{p+h} = \beta_0 + \beta_1 F1_p + \beta_2 F2_p \qquad
$$

The difference between `conditional_forc()` and `oos_vintage_forc()` is that in
the `conditional_forc()` function coefficients are only estimated once over all 
observations. Coefficients do not update based on what information would have been
available to the forecaster at any given point in time. We introduce stylized 
forecasts of X1 and X2 to demonstrate the `conditional_forc()` function. Because 
in this example we are making a conditional forecast for the future instead of testing past forecasts, we can condition on a horizon of forecasts. This is in contrast 
to the `oos_vintage_forc()` example where we test the performance of four quarter
ahead vintage forecasts.

```{r}
x1_forecast <- Forecast(
   origin   = as.Date(c("2012-06-30", "2012-06-30", "2012-06-30", "2012-06-30")),
   future   = as.Date(c("2012-09-30", "2012-12-31", "2013-03-31", "2013-06-30")),
   forecast = c(4.14, 4.04, 4.97, 5.12),
   realized = NULL,
   h_ahead  = NULL
)

x2_forecast <- Forecast(
   origin   = as.Date(c("2012-06-30", "2012-06-30", "2012-06-30", "2012-06-30")),
   future   = as.Date(c("2012-09-30", "2012-12-31", "2013-03-31", "2013-06-30")),
   forecast = c(6.01, 6.05, 6.55, 7.45),
   realized = NULL,
   h_ahead  = NULL
)

conditional_forc(
  lm_call = lm(y ~ x1 + x2, data),
  time_vec = data$date,
  x1_forecast, x2_forecast
)
```

This function is used to create a forecast for the present period or replicate 
a forecast made at a specific period in the past. Note that because we are 
forecasting into the future, `realized` is `NULL`. Also, because we are 
forecasting a horizon of dates, `h_ahead` is `NULL`.

## mse_weighted_forc

The `mse_weighted_forc()` function produces a weighted average of multiple 
forecasts based on the recent performance of each forecast. The function takes
two or more forecasts of the `Forecast` class, an evaluation window, and an error function. For each forecast period, the error function is used to calculate 
forecast accuracy over the past `eval_window` number of periods. The forecast accuracy of each forecast is then used to weight forecasts based on a weighting function. In each period, weights are calculated and used to create a weighted average forecast. We 
use a stylized example in which we create a weighted forecast of two forecasts: 
Y1 and Y2. 

For all periods *p* in the *k* number of forecasts *Y*, weights *W* are calculated over the `eval_window` *e* as:

$$
W_n = \frac{1/MSE(Y_{kE})}{1/\sum_{k=1}^{n}MSE(Y_{kE})} \qquad \text{where} 
\quad E = \{(p-e)..p\} 
$$

Forecasts are estimated as:

$$
forecast_p = Y1_pW_1 + Y2_pW_2
$$

```{r}
y1_forecast <- Forecast(
  origin = as.Date(c("2009-03-31", "2009-06-30", "2009-09-30", "2009-12-31",
                     "2010-03-31", "2010-06-30", "2010-09-30", "2010-12-31", 
                     "2011-03-31", "2011-06-30")),
  future = as.Date(c("2010-03-31", "2010-06-30", "2010-09-30", "2010-12-31",
                     "2011-03-31", "2011-06-30", "2011-09-30", "2011-12-31", 
                     "2012-03-31", "2012-06-30")),
  forecast = c(1.33, 1.36, 1.38, 1.68, 1.60, 1.55, 1.32, 1.22, 1.08, 0.88),
  realized = c(1.09, 1.71, 1.09, 2.46, 1.78, 1.35, 2.89, 2.11, 2.97, 0.99),
  h_ahead = 4L
)

y2_forecast <- Forecast(
  origin = as.Date(c("2009-03-31", "2009-06-30", "2009-09-30", "2009-12-31",
                     "2010-03-31", "2010-06-30", "2010-09-30", "2010-12-31", 
                     "2011-03-31", "2011-06-30")),
  future = as.Date(c("2010-03-31", "2010-06-30", "2010-09-30", "2010-12-31",
                     "2011-03-31", "2011-06-30", "2011-09-30", "2011-12-31", 
                     "2012-03-31", "2012-06-30")),
  forecast = c(0.70, 0.88, 1.03, 1.05, 1.01, 0.82, 0.95, 1.09, 1.07, 1.06),
  realized = c(1.09, 1.71, 1.09, 2.46, 1.78, 1.35, 2.89, 2.11, 2.97, 0.99),
  h_ahead = 4L
)

mse_weighted_forc(
  y1_forecast, y2_forecast,
  eval_window = 2L,
  errors = "mse",
  return_weights = FALSE
)
```

The `mse_weighted_forc()` function returns a weighted forecast of the Y1 and Y2
forecasts based on performance in recent periods. The weights used  in each 
period can be returned to the Global Environment by setting 
`return_weights` to `TRUE`. Note that although we were only weighting 
performance over the past two periods, we have five `NA` forecasts. This 
reflects the *lmForc* philosophy of replicating what it would be like to
forecast in real time. If a forecaster was making a forecast at `2010-06-30`,
they would only have access to realized values up to `2010-06-30`, in this case
the first two rows. This is why a weighted forecast with an `eval_window` of two
can only be computed once the forecast origin becomes `2010-06-30` and the 
forecaster has access to two realized values. This function can be used to 
compute a weighted forecast for the present period or to test how a weighted 
forecast would have performed historically.

</br>
</br>
